---
title: "Assignment #4"
author: "Julia Buschmann | jrb361"
date: "2/26/2022"
output: pdf_document
urlcolor: blue
source: https://usa.ipums.org/usa-action/data_requests/download
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

```{r include = FALSE}
library(tidyverse)
library(haven)
library(ipumsr)
library(ggplot2)
library(dplyr)
library(srvyr)
library(scales)
library(grid)
theme_set(theme_classic())
```


```{r, include = FALSE, results = 'hide'}
ddi <- read_ipums_ddi("cps_00002.xml")
data <- read_ipums_micro(ddi)
```


### Visualization #1
The below graph depicts the enrollment in Medicaid, differentiated by sex, for years 2016 through 2021. In each year, more women than men are enrolled in Medicaid, which could occur for a few reasons. Many states have family planning "demonstrations" which create a new eligibility group so women that are above the normal income limit for Medicaid can access contraception, exams, and fertility preservation services. Further, all states allow for income-eligible pregnant women to enroll in Medicaid (though their income limits differ dramatically, from 30% FPL to 138% FPL). Moreover, women typically work lower paying jobs than men, meaning that there are more likely to be eligible for Medicaid in all states. 

```{r}
data %>%
  group_by(YEAR) %>%
  filter (HIMCAIDLY == 2) %>%
  mutate(
    Sex = case_when(
      SEX == 1 ~ "Male",
      SEX == 2 ~ "Female"
    ))%>%
  ggplot +
  geom_bar(mapping = aes(x = factor(YEAR), fill = factor(Sex)), position = "dodge") +
  labs(title="Medicaid Enrollment 2016 to 2021, Differentiated By Sex", 
       caption="Source: IPUMS Micro Data",
       x = "Year",
       y = "Number of People Enrolled",
       fill = "Sex")
```

\newpage

### Visualization #2
A histogram demonstrating the age of Medicaid enrollees. Unsurprisingly, we see that the majority of enrollees are younger than 18. More than half of all births in the US are among Medicaid enrollees, and in some states, kids at certain income levels are eligible for various programs (e.g., the Children's Health Insurance Program) when their parents may not be.
```{r}
data %>%
  filter(YEAR == 2021) %>%
  filter(AGE <= 79)%>%
  filter(HIMCAIDLY == 2) %>%
  ggplot +
  geom_histogram(aes(x = AGE), 
                 binwidth = .1)+
  labs(title="Number of individuals enrolled in Medicaid, by Age, in 2021", 
       subtitle = "Due to coding methods, an accurate count of individuals 80 years and older is not available. 
       They have been excluded from the below graph.",
       caption="Source: IPUMS Micro Data",
       x = "Age of Medicaid Enrollee",
       y = "Number of People Enrolled (thousands)")+
  scale_x_continuous(breaks = breaks_width(10))

```

\newpage 

### Visualization #3
The below visualization demonstrates enrollment in private insurance, and the purchase of marketplace plans, varies by age. Unsurprisingly, enrollment in employer-sponsored insurance tracks closely with enrollment in private insurance. The difference between the two indicates that individuals are purchasing insurance from other sources, such as the health insurance marketplace. Here, we see a decline in private coverage and ESI around age 18, which could indicate that young adults are no longer covered under their parents' health insurance, which then picks up around age 25. We also see a steep decline at age 65, as people typically become eligible for Medicare at that age and may discontinue their ESI policy.

```{r}
data %>%
  filter(YEAR == 2021)%>%
  filter(PRVTCOVNW == 2 | GRPCOVNW == 2)%>%
  filter(AGE <= 79)%>%
  group_by(AGE)%>%
  summarize(PRVTCOVNW = sum(PRVTCOVNW), GRPCOVNW = sum(GRPCOVNW))%>%
  ggplot +
  geom_col(mapping = aes(x = AGE, y = PRVTCOVNW, fill = "skyblue"))+
  labs(title = "Enrollment in Private Health Insurance and Employer-Sponsored Insurance, 
       by Age, in 2021", 
       subtitle = "The black points indicate the number of individuals that have employer sponsored health inurance.",
       caption="Source: IPUMS Micro Data",
       x = "Age of Policy Holder",
       y = "Number of Individuals (thousands)") +
  theme(legend.position = "none")+
  geom_point(mapping = aes(x = AGE, y = GRPCOVNW, fill = "GRPCOVNW"))+
  scale_x_continuous(breaks = breaks_width(10))

```

\newpage

### Visualization #4
The below visualization shows the the number of experienced workers that were unemployed during 2021, distinguished by race. The uptick that occurred in March of 2021 could represent a sharp increase in unemployment as a product of the American Rescue Plan that was passed in February 2021, which included extended unemployment benefits. However, this sharp incline more likely occurred because of disaggregated survey results and the inclusion of an annual survey that takes place in March. (Note: I attempted to exclude the annual March survey, but all data points were affected.)

```{r}
data%>%
  filter(EMPSTAT == 21)%>%
  filter(YEAR == 2021)%>%
  mutate(
    month = case_when(
      MONTH == 01 ~ "January",
      MONTH == 02 ~ "February",
      MONTH == 03 ~ "March", 
      MONTH == 04 ~ "April",
      MONTH == 05 ~ "May",
      MONTH == 06 ~ "June",
      MONTH == 07 ~ "July",
      MONTH == 08 ~ "August", 
      MONTH == 09 ~ "September", 
      MONTH == 10 ~ "October", 
      MONTH == 11 ~ "November",
      MONTH == 12 ~ "December"
    ))%>%
  mutate(month = ordered(month, levels = unique(month))) %>%
  mutate(
    Race = case_when(
      RACE == 100 ~ "White",
      RACE == 200 ~ "Black",
      RACE == 300 ~ "American Indian/Alaska Native",
      RACE == 650 ~ "Asian/Pacific Islander",
      RACE == 651 ~ "Asian",
      RACE >= 801 & RACE <= 830 ~ "Two or More Races"
    ))%>%
  group_by(month, Race)%>%
  summarize(n = n())%>%
  ggplot +
  geom_point(mapping = aes(x = month, y = n, fill = Race, color = Race))+
  labs(title = "The Demographics of Unemployment in 2021", 
       subtitle = "The below graph shows the unemployment rate of experienced workers. New workers 
       also experienced a high rate of unemployment, with a similar demographic trend as 
       experienced workers.",
       x = "Month",
       y = "Number of Individuals (thousands)", 
       caption = "Source: IPUMS Micro Data")+
  geom_smooth(method = lm, se = FALSE, aes(x = month, y = n, color = Race))+
  geom_line(mapping = aes(x = month, y = n, fill = Race, color = Race))+
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1))+
  scale_y_continuous(breaks = breaks_width(1000))
```

